\bf Looking at Non-Linear Dimension Reductions as Models in the Data Space

### 2024-10-07 paper.qmd
Added best fit with tNSE
Updated Figure 12 Caption

```{r}
#| fig-cap: "Assessing which of the 5 NLDR layouts on the two_curvy_clust digit 1 data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e). Layout c is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-two_curvy_clust-mse
#| fig-pos: H
#| out-height: 60%

@@

distance_df_small_edges_two_curvy4 <- distance_two_curvy4 |>
  filter(distance < benchmark_two_curvy4) ## 0.221

@@

```{r}
#| label: best-fit-umap-tsne
#| fig-cap: "Best fit using UMAP ($a_1$ = 0.09) and tSNE ($a_1$ = 0.16) with the same number of non-empty bins ($m$ = 39). Both models show twist in $7\\text{-}D$."
#| fig-width: 10
#| fig-height: 20
#| out-height: 90%

@@

It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern. Interactivity also helps in understanding the quirks that occur with different NLDR techniques (XXXXRefer the video after finalising the S-curve).

>>>>

```{r}
#| fig-cap: "Assessing which of the 5 NLDR layouts on the two-curvy clusters data is the better representation using MSE for varying binwidth ($a_1$). Colour used for the lines and points in the left plot and in the scatterplots represents NLDR layout (a-e). Layout d is universally poor. Layouts a that show two close clusters are universally suboptimal. Layout a is the best choice."
#| label: fig-two_curvy_clust-mse
#| fig-pos: H
#| out-height: 60%

@@

distance_df_small_edges_two_curvy4 <- distance_two_curvy4 |>
  filter(distance < benchmark_two_curvy4) |> ## 0.221
  filter(!(row_number() %in% c(35, 89, 83, 25, 80, 56, 13, 59, 16)))

@@

```{r}
#| label: best-fit-umap-tsne
#| fig-cap: "Best fit for two-curvy clusters data using UMAP (a1-a4) ($a_1$ = 0.09) and tSNE (b1-b4) ($a_1$ = 0.16) with the same number of non-empty bins ($m$ = 39). Both models show twist in $7\\text{-}D$ (a3, a4, b3, b4). Also, both NLDR methods are unable to accurately capture the width of the curvilinear clusters."
#| fig-width: 10
#| fig-height: 20
#| out-height: 90%

@@

It’s important to access the \gD{} layout and the generated model overlaid on data in \pD{} together to understand whether it fits the points everywhere, fits better in some places, or simply mismatches the pattern. Interactivity also helps in understanding the quirks that occur with different NLDR techniques (XXXXRefer the video after finalising the two-curvy clusters data).

### 2024-10-07 paper.tex
Updated Figure 10
Updated Figure 12 Caption

\caption{Distribution of edge lengths and wireframes in \gD{} and
\(7\text{-}D\) with Delaunay triangulation, and two choices of long edge
removal: (b) benchmark = 2\(a_1\) and (c) benchmark = 5\(a_1\), where
\(a_1\) = 0.09.}

@@

\caption{\label{fig-two_curvy_clust-mse}Assessing which of the 5 NLDR
layouts on the two\_curvy\_clust digit 1 data is the better
representation using MSE for varying binwidth (\(a_1\)). Colour used for
the lines and points in the left plot and in the scatterplots represents
NLDR layout (a-e). Layout c is universally poor. Layouts a that show two
close clusters are universally suboptimal. Layout a is the best choice.}

@@

\caption{Best fit using UMAP (\(a_1\) = 0.09) and tSNE (\(a_1\) = 0.16)
with the same number of non-empty bins (\(m\) = 39). Both models show
twist in \(7\text{-}D\).}

@@

It's important to access the \gD{} layout and the generated model
overlaid on data in \pD{} together to understand whether it fits the
points everywhere, fits better in some places, or simply mismatches the
pattern. Interactivity also helps in understanding the quirks that occur
with different NLDR techniques (XXXXRefer the video after finalising the
S-curve).

>>>>

\caption{Distribution of edge lengths and wireframes in \gD{} and
\(7\text{-}D\) with Delaunay triangulation, and two choices of long edge
removal: (b) benchmark = 4.5\(a_1\) and (c) benchmark = 2\(a_1\), where
\(a_1\) = 0.09.}

@@

\caption{\label{fig-two_curvy_clust-mse}Assessing which of the 5 NLDR
layouts on the two-curvy clusters data is the better representation
using MSE for varying binwidth (\(a_1\)). Colour used for the lines and
points in the left plot and in the scatterplots represents NLDR layout
(a-e). Layout d is universally poor. Layouts a that show two close
clusters are universally suboptimal. Layout a is the best choice.}

@@

\caption{Best fit for two-curvy clusters data using UMAP (a1-a4)
(\(a_1\) = 0.09) and tSNE (b1-b4) (\(a_1\) = 0.16) with the same number
of non-empty bins (\(m\) = 39). Both models show twist in \(7\text{-}D\)
(a3, a4, b3, b4). Also, both NLDR methods are unable to accurately
capture the width of the curvilinear clusters.}

@@

It's important to access the \gD{} layout and the generated model
overlaid on data in \pD{} together to understand whether it fits the
points everywhere, fits better in some places, or simply mismatches the
pattern. Interactivity also helps in understanding the quirks that occur
with different NLDR techniques (XXXXRefer the video after finalising the
two-curvy clusters data).
